\newsection{Shares}
\index{Share}
\index{Dividend}
\label{shares}
A~share is a unit of~a~company's \hyperref[capital]{capital} owned by~someone else.
This someone, who~owns the~unit, is called \textit{shareholder}.
Simply said a~shareholder owns a~small part of~a~company.
By~buying a~share the~shareholder gives his money to~increase the~\hyperref[capital]{capital} of~the~company.
And~the~question is obvious.
Why~would anyone give his money to~a~company capital?

Companies motivate shareholders to~buy its shares by so~called \textit{dividends}.
A~dividend is an~amount of~money paid to~shareholders, typically once per~year.
When the~company's business has~been successful for~last year, the~company uses some part of~the~total profit for~dividends.
Each~shareholder will get a~portion of~this profit part -- a~dividend -- based on~the~number of~shares he~has.
If~the~dividend is higher than the~money he~had~paid for~his shares, then it~was a~successful share trade.

\newsubsection{Share VS Stock}
\index{Stock}
\index{Stockholder}
Share and~stock are very close synonyms.
In~modern English the~plural \itq{shares} is used almost interchangeably with \itq{stock}.
Also, the~term \itq{stockholder} is~used for~someone, who~owns stock.

A~slight difference between these two terms exists in~the~level of~generality.
When talking about a~concrete company, shares are used.
For~example:
"I'm~a~shareholder, I~own shares of~Microsoft."
When talking in~general, stock is used.
For~example:
"I'm~a~stockholder, I~own stock."
