\newsection{Stock Market}
\index{Stock market}
\hyperref[shares]{Dividends} aren't the~only way how~to~get money by \hyperref[shares]{shares} ownership.
The~other way, and~actually much more known and~profitable, is~(buying~and)~selling owned shares in~stock markets.
When a~share has some price in~a~stock market, it~means that there's some idiot willing to~pay the~price for~the~share and~hoping to~get more from dividends or~to~sell it for~even more to~another idiot.
That's the~whole mechanism of~stock markets, there's really nothing else.
Examples of~stock market idiots are crazy individuals, banks, all~sorts of~financial and~investment groups, mafia~etc.

In~today's automated world selling stock may~seem slightly mysterious, because sellers usually don't~know who~buyers~are.
Buyers have pre--registered offers to~buy shares and~pay certain amount for~them in~the~market system.
So~when~you~sell shares, you~just press a~big \mbit{Sell} button in~the~system and~all~transactions are~processed automatically.
You~will~never know who~that~idiot~was.

There are many stock markets.
Each~has specific conditions for~companies and~shareholders to~fulfil to~be~allowed to~do share transactions.
The~bigger the~market is, the~better chance to~gain \hyperref[capital]{capital} for~companies and~to~earn money for~shareholders.
But~of~course, bigger markets usually have stricter conditions for~participants.
Many countries have a~big stock market named after the~capital city, like the~stock market of~London, Paris or~Beijing.
The~biggest and~most important stock market in~the~World is, of~course, the~one in~Wall~Street in~New~York.
\newpage
